FDO-based agent systems require more than execution traces when a single operation must be checked outside the original runtime context. They need a compact accountability statement that can specify which operation was performed on which digital object context, under which policy basis, with which input and output references, what evidence was produced, and how the statement can be independently validated. This paper presents a minimal verifiable profile for that purpose and grounds it in a current reference implementation. The profile is intentionally narrow: it centers operation accountability and binds \texttt{operation}, \texttt{policy}, \texttt{provenance}, \texttt{evidence}, and \texttt{validation} into one portable, externally checkable statement while keeping only the supporting context needed for independent review. The implementation provides a profile specification, a JSON Schema, two valid examples, five invalid examples, a profile-aware validator, a command-line entry point, tests, and a runnable single-path demo. The validator goes beyond structural schema checks to cover reference closure, cross-field consistency, and minimal integrity checks, and it produces machine-readable JSON reports, human-readable summaries, and explicit error codes. The review package also preserves a frozen artifact path, while public release and archive identifiers are withheld for anonymous review. The paper does not claim broad cross-framework validation or industrial-scale deployment. Its claim is narrower: in the current implementation context, operation accountability can be packaged as a minimal, testable, reproducible, and archivable engineering unit.
