% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[
]{article}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\newenvironment{Shaded}{}{}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.49,0.56,0.16}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.53,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.56,0.13,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.73,0.13,0.13}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{1.00,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.25,0.63,0.44}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.02,0.16,0.49}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.50,0.00}{\textbf{#1}}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.40,0.40,0.40}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.44,0.13}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.74,0.48,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.73,0.40,0.53}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.10,0.09,0.49}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.25,0.44,0.63}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.38,0.63,0.69}{\textbf{\textit{#1}}}}
\usepackage{longtable,booktabs,array}
\newcounter{none} % for unnumbered tables
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{}

\begin{document}

\section{AEP-Media: Reusable Research Software for Offline Validation of
Time-Aware Media Evidence
Bundles}\label{aep-media-reusable-research-software-for-offline-validation-of-time-aware-media-evidence-bundles}

\subsection{Authors}\label{authors}

Bin Zhang

Independent Researcher

ORCID: 0009-0002-8861-1481

Email: joy7759@gmail.com

\subsection{Abstract}\label{abstract}

AEP-Media is reusable research software for representing a media-bearing
operation as a locally checkable evidence bundle. It provides a minimal
time-aware media evidence profile, JSON schemas, controlled examples,
validators, command-line tools, offline bundle build and verification,
strict declared time-trace validation, adapter-only ingestion
interfaces, evaluation matrices, and release packaging reports. The
software is intended for researchers studying operation accountability,
media evidence packaging, provenance boundaries, and reproducible
validation artifacts. AEP-Media checks profile identity, required
fields, reference closure, artifact hashes, path safety, bundle
checksums, declared clock-trace coverage, clock-trace summaries, and
offset/jitter thresholds. Its evaluation materials exercise valid,
invalid, tamper, adapter-ingestion, and optional-tool reporting cases
with expected pass/fail behavior. The current claim is local validation
and fixture-based adapter ingestion. AEP-Media does not claim legal
admissibility, non-repudiation, trusted timestamping, real PTP proof,
full MP4 PRFT parsing, real C2PA signature verification, chain of
custody, or production deployment.

\subsection{Keywords}\label{keywords}

media evidence; operation accountability; research software; provenance;
offline validation; evidence bundle

\subsection{Metadata}\label{metadata}

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Nr
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Code metadata description
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Metadata
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
C1 & Current code version & \texttt{aep-media-v0.1.0} \\
C2 & Permanent link to code/repository used for this code version &
\texttt{https://github.com/joy7758/agent-evidence} \\
C3 & Legal code license & Apache-2.0 \\
C4 & Code versioning system used & git \\
C5 & Software code languages, tools and services used & Python; JSON
Schema; pytest; Click CLI; Pandoc/LaTeX only for manuscript packaging \\
C6 & Compilation requirements, operating environments and dependencies &
Python 3.11+; install with \texttt{python\ -m\ pip\ install\ -e\ .};
optional LinuxPTP, FFmpeg/ffprobe, and C2PA tools are not required for
fixture-based validation \\
C7 & If available, link to developer documentation/manual &
\texttt{https://github.com/joy7758/agent-evidence}; repository
documentation under \texttt{README.md}, \texttt{spec/},
\texttt{schema/}, \texttt{examples/media/}, \texttt{demo/}, and
\texttt{docs/reports/} \\
C8 & Support email for questions & joy7759@gmail.com \\
\end{longtable}
}

\subsection{1. Motivation and
significance}\label{motivation-and-significance}

Researchers working with media-bearing operations often need to review
disconnected artifacts: media files, logs, provenance declarations,
timing metadata, policy references, and validation reports. These
objects are difficult to inspect consistently when they are not bound
into a local evidence object with explicit validation semantics.
AEP-Media addresses this gap by providing a reusable software layer for
turning such materials into locally checkable evidence bundles with
machine-readable failure codes.

The problem is deliberately bounded. AEP-Media is not a legal evidence
platform and does not attempt to prove external authenticity. Its
purpose is to make the local validation boundary explicit: which fields
must be present, which references must close, which files must match
declared hashes, which time traces cover a declared window, and which
failures are reported when a controlled rule is broken. This local
conformance layer is useful before stronger external assurance
mechanisms such as signing, trusted timestamping, or deployed capture
pipelines are added.

AEP-Media is positioned next to provenance, media-authenticity, timing,
and artifact-review surfaces rather than as a replacement for them. W3C
PROV provides general provenance concepts {[}1,2{]}. C2PA focuses on
content provenance and manifest-based authenticity {[}3{]}.
FFmpeg/ffprobe and LinuxPTP expose media timing and
clock-synchronization surfaces {[}4,5{]}, while GStreamer provides a PTP
clock interface for media pipelines {[}6{]}. JSON Schema supports
structured validation {[}7{]}, and artifact-review practices emphasize
inspectable and reproducible software packages {[}8{]}. AEP-Media
contributes a concrete local validation boundary that binds media
artifacts, declared provenance, declared time traces, bundle integrity,
and failure codes into one reusable research software artifact.

\subsection{2. Software description}\label{software-description}

AEP-Media extends the \texttt{agent-evidence} package with a
media-focused profile and validation path. A media evidence statement
binds an actor, subject, operation, policy, constraints, time context,
media artifacts, provenance references, evidence notes, and validation
expectations into one object. The profile validator checks that the
statement is structurally complete and locally consistent.

The offline bundle layer packages a statement and its artifacts into a
portable directory. The builder copies artifacts, rewrites paths to
bundle-local locations, recomputes hashes and sizes, writes checksums,
and emits validation reports. The verifier rejects unsafe paths, missing
artifacts, checksum mismatches, and media profile failures. This allows
review without relying on the original runtime layout.

The strict-time layer validates declared clock-trace artifacts. It
checks trace references, artifact roles, trace hashes, trace profile
identity, collection window coverage, sample validity, summary
recomputation, offset thresholds, jitter thresholds, and media binding
to the declared time context.

The adapter layer is ingestion-only. LinuxPTP-style logs, FFmpeg
PRFT-style timing metadata, and C2PA-like manifest metadata can be
normalized into AEP-Media reports. Adapter reports distinguish fixture
ingestion from optional external-tool reporting and do not imply
external verification unless a future environment-specific run records
it. This boundary is important because real clock discipline, PRFT
extraction, and C2PA signature verification depend on external tools and
environment-specific evidence.

The main modules are:

\begin{itemize}
\tightlist
\item
  \texttt{agent\_evidence/media\_profile.py}: media profile validation.
\item
  \texttt{agent\_evidence/media\_bundle.py}: offline bundle build and
  verification.
\item
  \texttt{agent\_evidence/media\_time.py}: strict declared time-trace
  validation.
\item
  \texttt{agent\_evidence/adapters/linuxptp.py}: LinuxPTP-style trace
  ingestion.
\item
  \texttt{agent\_evidence/adapters/ffmpeg\_prft.py}: FFmpeg PRFT-style
  metadata ingestion.
\item
  \texttt{agent\_evidence/adapters/c2pa\_manifest.py}: C2PA-like
  manifest ingestion.
\item
  \texttt{agent\_evidence/media\_evaluation.py}: default evaluation
  matrix.
\item
  \texttt{agent\_evidence/media\_adapter\_evaluation.py}:
  adapter-inclusive evaluation.
\item
  \texttt{agent\_evidence/media\_optional\_tools.py}: optional
  external-tool reporting.
\item
  \texttt{agent\_evidence/media\_release\_pack.py}: release pack
  assembly.
\end{itemize}

\subsection{3. Illustrative examples}\label{illustrative-examples}

A typical profile validation run is:

\begin{Shaded}
\begin{Highlighting}[]
\ExtensionTok{agent{-}evidence}\NormalTok{ validate{-}media{-}profile }\DataTypeTok{\textbackslash{}}
\NormalTok{  examples/media/minimal{-}valid{-}media{-}evidence.json}
\end{Highlighting}
\end{Shaded}

A minimal offline bundle workflow is:

\begin{Shaded}
\begin{Highlighting}[]
\ExtensionTok{agent{-}evidence}\NormalTok{ build{-}media{-}bundle }\DataTypeTok{\textbackslash{}}
\NormalTok{  examples/media/minimal{-}valid{-}media{-}evidence.json }\DataTypeTok{\textbackslash{}}
  \AttributeTok{{-}{-}out}\NormalTok{ /tmp/aep{-}media{-}bundle{-}check}

\ExtensionTok{agent{-}evidence}\NormalTok{ verify{-}media{-}bundle }\DataTypeTok{\textbackslash{}}
\NormalTok{  /tmp/aep{-}media{-}bundle{-}check}
\end{Highlighting}
\end{Shaded}

A strict-time validation run is:

\begin{Shaded}
\begin{Highlighting}[]
\ExtensionTok{agent{-}evidence}\NormalTok{ validate{-}media{-}time{-}profile }\DataTypeTok{\textbackslash{}}
\NormalTok{  examples/media/time/minimal{-}valid{-}time{-}aware{-}media{-}evidence.json}
\end{Highlighting}
\end{Shaded}

The evaluation matrix can be reproduced with:

\begin{Shaded}
\begin{Highlighting}[]
\ExtensionTok{agent{-}evidence}\NormalTok{ run{-}media{-}evaluation }\DataTypeTok{\textbackslash{}}
  \AttributeTok{{-}{-}out}\NormalTok{ /tmp/aep{-}media{-}evaluation}
\end{Highlighting}
\end{Shaded}

The repository includes valid examples, controlled invalid examples,
strict-time fixtures, adapter fixtures, and tamper cases. These allow
researchers to reproduce the intended pass/fail behavior without
installing LinuxPTP, FFmpeg, ffprobe, or C2PA. The archived software
release is available from Zenodo as version \texttt{aep-media-v0.1.0}
{[}9{]}.

\subsection{4. Impact}\label{impact}

AEP-Media supports research on operation accountability, media evidence
packaging, provenance boundaries, and reproducible validation artifacts.
It gives researchers a concrete software artifact for asking whether a
declared media evidence package is internally consistent and portable.
This is useful for experiments that need a stable evidence-object
representation rather than ad hoc media files and logs.

The evaluation is conformance-oriented. Prior release reports record 18
default cases, 26 adapter-inclusive cases, 23 optional-tool reporting
cases, and 31 combined adapter/optional-tool cases, all with
\texttt{unexpected=0}. The case categories include valid conformance
cases, invalid single-rule cases, bundle tamper cases, strict-time
failures, adapter-ingestion cases, and optional-tool reporting cases.
Expected failures include missing time context, media hash mismatch,
unresolved policy reference, bundle checksum mismatch, path escape,
missing clock-trace reference, clock offset threshold exceedance, clock
window mismatch, missing PRFT metadata, and declared invalid C2PA-like
signature status.

The impact is not that AEP-Media proves authenticity. Its value is that
it makes the local evidence object explicit and reproducibly checkable,
which is a prerequisite for stronger assurance layers. Researchers can
use it as a baseline when comparing provenance models, evidence-package
formats, adapter boundaries, or future external-verification appendices.
The implementation also complements supply-chain and attestation systems
such as in-toto and SLSA by focusing on media-bearing operation evidence
rather than build provenance alone {[}10,11{]}.

\subsection{5. Conclusions}\label{conclusions}

AEP-Media is reusable research software for offline validation of
time-aware media evidence bundles. It combines a small profile, schemas,
examples, command-line validators, bundle verification, strict declared
time-trace validation, adapter-only ingestion, reproducible evaluation
matrices, and release packaging material.

The current claim is local validation and fixture-based adapter
ingestion. AEP-Media detects inconsistencies inside a declared evidence
package, but it does not establish that the original media capture event
was truthful, authorized, or unmodified before packaging. It does not
provide legal admissibility, non-repudiation, trusted timestamping, real
PTP proof, full MP4 PRFT parsing, real C2PA signature verification,
chain of custody, production deployment, or broad forensic sufficiency.

Future work should keep the local profile semantics stable while adding
environment-specific evidence appendices: LinuxPTP traces captured on
equipped Linux hosts, ffprobe output from PRFT-bearing media, C2PA CLI
reports for real signed assets, and independent reproduction of the
evaluation matrix.

\subsection{Data availability}\label{data-availability}

The software, selected examples, schemas, reports, and release materials
are available in the public GitHub repository at
\texttt{https://github.com/joy7758/agent-evidence} and in the Zenodo
archive for AEP-Media v0.1.0 at
\texttt{https://doi.org/10.5281/zenodo.20107097}. The supplementary
package accompanying this submission contains selected schemas,
examples, reproducibility commands, evaluation summaries, and
claim-boundary documentation.

\subsection{CRediT authorship contribution
statement}\label{credit-authorship-contribution-statement}

Bin Zhang: Conceptualization, Methodology, Software, Validation,
Investigation, Data curation, Writing - original draft, Writing - review
and editing, Project administration.

\subsection{Declaration of competing
interest}\label{declaration-of-competing-interest}

The author declares no known competing financial interests or personal
relationships that could have appeared to influence the work reported in
this article.

\subsection{Funding}\label{funding}

No specific funding was received for this work.

\subsection{Declaration of generative AI and AI-assisted technologies in
the manuscript preparation
process}\label{declaration-of-generative-ai-and-ai-assisted-technologies-in-the-manuscript-preparation-process}

During the preparation of this work, the author used OpenAI
ChatGPT/Codex for manuscript organization, implementation scaffolding,
command generation, and wording refinement. After using these tools, the
author reviewed and edited the content as needed and takes full
responsibility for the content of the submitted article.

\subsection{References}\label{references}

{[}1{]} W3C. PROV Overview. World Wide Web Consortium.
\url{https://www.w3.org/TR/prov-overview/}; 2013 {[}accessed 10 May
2026{]}.

{[}2{]} W3C. PROV-DM: The PROV Data Model. World Wide Web Consortium.
\url{https://www.w3.org/TR/prov-dm/}; 2013 {[}accessed 10 May 2026{]}.

{[}3{]} Coalition for Content Provenance and Authenticity. C2PA
Technical Specification. \url{https://c2pa.org/specifications/}; 2024
{[}accessed 10 May 2026{]}.

{[}4{]} FFmpeg Project. ffprobe Documentation.
\url{https://ffmpeg.org/ffprobe.html}; 2026 {[}accessed 10 May 2026{]}.

{[}5{]} LinuxPTP Project. ptp4l and phc2sys Documentation.
\url{https://linuxptp.sourceforge.net/}; 2026 {[}accessed 10 May
2026{]}.

{[}6{]} GStreamer Project. GstPtpClock Documentation.
\url{https://gstreamer.freedesktop.org/documentation/net/gstptpclock.html};
2026 {[}accessed 10 May 2026{]}.

{[}7{]} JSON Schema. JSON Schema Core and Validation Specifications.
\url{https://json-schema.org/specification}; 2026 {[}accessed 10 May
2026{]}.

{[}8{]} Association for Computing Machinery. Artifact Review and
Badging.
\url{https://www.acm.org/publications/policies/artifact-review-and-badging-current};
2026 {[}accessed 10 May 2026{]}.

{[}9{]} Zhang B. AEP-Media v0.1.0: Offline Validation of Time-Aware
Media Evidence Bundles {[}software{]}. Zenodo; 2026.
\url{https://doi.org/10.5281/zenodo.20107097}.

{[}10{]} in-toto Project. in-toto: Providing farm-to-table guarantees
for bits and bytes. \url{https://in-toto.io/}; 2026 {[}accessed 10 May
2026{]}.

{[}11{]} Supply-chain Levels for Software Artifacts. SLSA Provenance.
\url{https://slsa.dev/spec/v1.0/provenance}; 2026 {[}accessed 10 May
2026{]}.

\end{document}
